3D seismic diffraction separation and imaging using the local rank-reduction method

### 2022-01-02 diffrlrra3d/paper.tex
Sun Jan  2 01:07:39 CST 2022

\documentclass[10pt]{IEEEtran}
\usepackage{amsmath}
\usepackage{cite}
\usepackage{ifthen}
\usepackage{seg}
\usepackage{color}
\usepackage{graphicx}
\usepackage{subfigure}

\DeclareRobustCommand{\dlo}[1]{}
\DeclareRobustCommand{\wen}[1]{#1}
\DeclareRobustCommand{\old}[1]{}
\DeclareRobustCommand{\new}[1]{#1}

\begin{document}

\title{3D seismic diffraction separation and imaging using the local rank-reduction method}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\author{Wei Chen, Xingye Liu, Omar M. Saad, Yapo Abol{\'e} Serge Innocent Obou{\'e}, Liuqing Yang, Hang Wang, and Yangkang Chen
\thanks{W. Chen is with Cooperative Innovation Center of Unconventional Oil and Gas (Ministry of Education \& Hubei Province), Yangtze University, Wuhan, 430100, China, and Key Laboratory of Exploration Technologies for Oil and Gas Resources of Ministry of Education, Yangtze University, Wuhan, 430100, China}
\thanks{X. Liu is with Chengdu University of Technology, Dongsanlu, Erxianqiao, Chengdu, 610059, China.}
\thanks{O.M. Saad is with Seismology Department, National Research Institute of Astronomy and Geophysics (NRIAG), Helwan, 11731, Egypt.}
\thanks{Y.A.S.I. Oboue and H. Wang are with School of Earth Sciences, Zhejiang University, Hangzhou, 310027, China.}
\thanks{L. Yang is with State Key Laboratory of Petroleum Resource and Prospecting, China University of Petroleum (Beijing), Beijing, 102200, China.}
\thanks{Y. Chen is with Bureau of Economic Geology, Jackson School of Geosciences, The University of Texas at Austin, Austin, Texas, 78713-8924, USA.}
\thanks{The research is supported by the National Natural Science Foundation of China (Grant No. 41804140) and the Open Fund of Cooperative Innovation Center of Unconventional Oil and Gas, Yangtze University (Ministry of Education \& Hubei Province) (Grant No. UOG2021-01).}}

@@

\begin{keywords}
Seismic signal separation, seismic denoising, seismic imaging, seismic diffraction
\end{keywords}

\section{Introduction}
Seismic data contain various useful information that can reflect the geologic features of the subsurface \cite{shucai2020,guoyin2018}. In general, reflections are generated by large-scale geologic objects that are continuous and have smooth-reflecting boundaries, and their energy in the seismic records is relatively strong \cite{2010Imaging, Peng2018Accurate}. 
Reflections are preserved in conventional seismic exploration workflow, and are the most commonly used signals to detect large-scale geologic structures. 
However, it is difficult to recognize small-scale geologic structures  (faults, small-size scattering points, and reflector edges, etc.) by using reflections. 
These small-scale geologic bodies will scatter the energy that is expressed as diffractions in the seismic records, which are destroyed or removed by the traditional seismic processing techniques.
However, diffractions are helpful to find unconventional reservoirs of oil and gas because they are associated with small-scale discontinuous objects \cite{2016Diffraction, 2020Separation}.

Researchers use seismic diffraction to detect subsurface discontinuities, provided that the seismic diffraction is separated from the original data.  
Nevertheless, it is a challenge because the energy of diffractions is weaker compared with the strong energy of reflections, which can even reach several orders of magnitude, i.e., diffractions are usually masked by reflections and noise \cite{1994Theory, 2007Post,omar2020geo1}. 
One of the most widely used methods is the path-summation-based method that uses the kinematic differences to separate reflections and diffractions. Based on the coherent summation of diffracted events according to the traveltime, \cite{1987A} and \cite{1988Imaging} achieve diffraction separation and imaging in common-offset and common-midpoint domains, respectively. 
\cite{2004Diffraction} separate and mute the reflection in the pseudodepth domain  by using the difference in properties of waves.
The most known diffraction separation and imaging method is based on a common-reflection-surface method that has been extended by many geophysicists.
 One of the advantages is that this method using several common midpoints to stack based on ray theory \cite{2013Recovering}.
\cite{2009Diffraction} propose a multi-focusing algorithm to focus diffractions while scattering reflections.
Similarly, \cite{2011Common} use the common-reflection-surface approach to achieve diffraction imaging.
\cite{2012Separation} achieve diffraction imaging in the migrated dip-angle domain and pointed out that it was more advantageous to separate diffraction from seismic data. \cite{2013Diffraction} propose a method that can adaptively choose parameters to fit the diffraction trvaltime and accelerate the computation.
\cite{Coimbra2018Enhancement} improve the accuracy of diffraction separation by a finite-offset double-square equation, and extended a fast version without losing accuracy.
\cite{2018Fast} develop a robust and fast method to extract the parameters of common-reflection-surface by using the local slope attribute, and it could be used for the diffraction separation. 
\cite{2018Common} point out that wavefront attributes of pre-stack seismic data could improve the illumination and are more appropriate for diffraction separation.
\cite{20193D} construct a new traveltime difference gather to flatten the diffraction, then muted the upward reflection.
For post-stack seismic data, \cite{2019Coherent} uses many wavefront filters to separate reflections and diffractions, and demonstrated its superiority in extracting diffractions masked by strong reflections.

Another effective diffraction separation and imaging method is based on the plane-wave destruction (PWD) proposed by \cite{Sergey2002Applications} and  \cite{2007Post}. Traveltimes of diffractions are quasi-hyperbolic-shaped while those of reflections are laterally continuous in the constant plane-wave slope sections. Thus, reflections can be predicted from neighbor traces by using a prediction-error filter. Then, the predicted reflection is suppressed and the diffraction is extracted.  After that, diffractions are migrated by selecting an optimal velocity. That is to say, the PWD-based method can simultaneously achieve velocity estimation and diffraction imaging \cite{2011Azimuthally}. 
\cite{2013Coal} utilize the PWD-based diffraction imaging method to detect the fractures in coals.
\cite{Binzhong2018seismic} propose a local moving-average-error-filter to achieve diffraction separation and applied it to coal structures detection. 
\cite{2018l1} developed an improved PWD-based method via wavelet transform and L1-norm, and achieved accurate diffraction separation.
The PWD-based method depends on the local slopes, thus a wide range of approaches are developed to accurately estimate the local slopes, such as \cite{2008Separation, 2010Nonlinear, 2018Dip}.
\cite{2020Least} achieve the diffraction separation by establishing an inversion objection and combining the PWD algorithm with other advanced techniques, such as Kirchhoff modeling, and path-summation-based methods.

Most diffraction separation and imaging methods only take kinematic features of diffractions and reflections into consideration, and the dynamic differences between diffractions and reflections are ignored. However, the dynamic discrepancy is significant to diffraction imaging \cite{2015Diffraction, 20193D, 2020Diffraction}.
\cite{2014Diffraction} update migration velocities according to shapes of reflections and diffractions in the migrated gathers by picking the angles related to reflections in the shot and opening-angle gathers, improving the accuracy.
\cite{2015Diffraction} develop a least-squares fitting method by using double exponential functions to analyze the amplitude function of diffracted waves and improve the performance of diffraction separation.
\cite{2016Imaging} develop a three-step diffraction separation and imaging approach based on wave-equation. It is computationally fast and resistant to random noise.
Based on the fact that the dip direction of reflectors could not be observed in both left- and right-downgoing propagating waves, \cite{2019Efficient} develop an efficient wave-equation-based approach for diffraction separation.
In frequency-wavenumber domain, \cite{2020Inspiration} use a double square root equation to simulate diffractions, and separated diffractions by using dip frequency filters.
To improve the image quality of diffractions, \cite{2020Periodic} propose a periodic plane-wave least-squares reverse time migration method.
\cite{2021Diffraction} applies a diffraction-angle filtering method to the virtual source in the acoustic full-waveform inversion to separate energy of reflections and transmissions. Diffractions can be separated by designing many types of diffraction-angle filters.

Consequently, both kinematic and dynamic differences between reflections and diffractions are important for seismic diffraction separation and imaging. Jointly using kinematic and dynamic properties of wave fields is helpful for diffraction separation and imaging \cite{2010Separation}. 
\cite{2020Separation} propose a 2D diffraction separation method based on a local rank-reduction algorithm, in which the rank is optimized by an adaptive selection approach.  \cite{2020Diffraction} utilize the multichannel singular-spectrum  analysis (MSSA) for diffraction separation and imaging. Their implementation of the MSSA method does not consider the strong non-stationarity of seismic data and thus does not use local windows for the diffraction separation, which is referred to as a global rank-reduction (GRR) method. In the post-stack domain, these two methods make full use of kinematic and dynamic differences between reflections and diffractions. Besides, these methods were tested on 2D seismic data. For 3D data, the complexity would increase and a unique rank value is unsatisfactory for accurate diffraction separation and imaging. 
Therefore, we present a novel method based on the local rank-reduction (LRR) method for 3D seismic diffraction separation. Both kinematic and dynamic properties are considered. The rank-reduction filter is carried out in each local window to enhance the performance \cite{shaohuan2017gji,yangkang2017ieee,yangkang2019nc}. In this case, specifying an appropriate rank for different windows is necessary to make the result more promising. Due to the rank inconsistency,  we propose an automatic selection approach to locally optimize the rank at different locations, which is not essential in the global strategy.

@@

\begin{figure}[htb!]
\centering
\includegraphics[width=\columnwidth]{Fig/demosigma}
\caption{Singular value distribution. (a) 3D reflection data containing three plane waves. (b) 3D seismic data containing reflections, diffractions and random noise. (c) 2D section extracted from (a). (d) 2D section extracted from (b). (e) Singular values distribution of the Hankel matrix in the 20 Hz slice of the data shown in (a).  (f) Singular values distribution of the data shown in (b). }
\label{fig:demosigma}
\end{figure}

\begin{figure}[htb!]
 \centering
 \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/refl}
   \label{fig:refl}}
 \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/diffr}
   \label{fig:diffr}}   
 \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/syn}
   \label{fig:syn}}    
\caption{Benchmark dataset with groundtruth reflections and diffractions. (a) Reflection data. (b) Diffraction data. (c) Synthetic data. }
\label{fig:refl,diffr,syn}
\end{figure}

\begin{figure}[htb!]
 \centering
 \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-pwd-0}
   \label{fig:s-pwd-0}}
 \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-grr-0}
   \label{fig:s-grr-0}}   
 \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-lrra-0}
   \label{fig:s-lrra-0}}    
\caption{Separated reflections using (a) PWD method, (b) global rank-reduction method, and (c) proposed method.}
\label{fig:s-pwd-0,s-grr-0,s-lrra-0}
\end{figure}

\begin{figure}[htb!]
 \centering
 \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-pwd-n-0}
   \label{fig:s-pwd-n-0}}
 \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-grr-n-0}
   \label{fig:s-grr-n-0}}   
 \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-lrra-n-0}
   \label{fig:s-lrra-n-0}}    
\caption{Separated diffractions using (a) PWD method (SNR=6.52 dB), (b) global rank-reduction method (SNR=1.42 dB), and (c) proposed method (SNR=9.14 dB).}
\label{fig:s-pwd-n-0,s-grr-n-0,s-lrra-n-0}
\end{figure}

\begin{figure}[htb!]
\centering
\subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-pwd-e-0}
  \label{fig:s-pwd-e-0}}
\subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-grr-e-0}
  \label{fig:s-grr-e-0}}   
\subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-lrra-e-0}
  \label{fig:s-lrra-e-0}}    
\caption{\new{Error of separated diffractions using (a) PWD method, (b) global rank-reduction method, and (c) proposed method.} \color{black}{}}
\label{fig:s-pwd-e-0,s-grr-e-0,s-lrra-e-0}
\end{figure}

\section*{THEORY AND Methods}
Diffraction separation can be regarded as a problem of seismic data reconstruction, in which diffractions are treated as effective signals while reflections are treated as noise.  In the post-stack seismic data, reflections appear as locally linear events, while diffractions behave as hyperbolic events.  The rank-reduction method can utilize the kinematic differences between diffractions and reflections to separate different components. Besides, the energy of diffractions is much weaker than that of reflections, thereby corresponding to different singular values and singular vectors. The distributions of singular values for diffractions and reflections are assumed to be disparate in the singular-value spectrum. Therefore, the contribution of the last few singular values is the smallest, while the first few singular values contribute more to reconstructing the data. The rank-reduction method also considers the energy discrepancy, i.e., the dynamic differences, to separate weak diffractions and strong reflections \cite{2010Separation, 2016Diffraction}.

\subsection{Local rank-reduction method}
Let a 3D seismic data be denoted by $S(x, y, t)$, $x=1,\cdots, N_x$, $y=1,\cdots, N_y$, and $t=1,\cdots, N_t$, where $N_x$, $N_y$ and $N_t$ represent the numbers of samples in spatial and time axes, respectively. The rank-reduction method \cite{2011Simultaneous,weilin2016dmssa} is implemented in the frequency domain, thus the first step is to obtain the corresponding frequency-domain data $S(x, y,f )$  by forward Fourier transform $\mathcal{F}$. For a given frequency $f_0$, we construct the block Hankel matrix $\mathbf{{H}}$ by mapping the frequency-domain data into several Hankel matrices $\mathbf{\widetilde{H}}_i(f_0)$ through the Hankelization operator $\mathcal{H}$. Then, we can obtain,
\begin{align} \label{eq:B} 
\mathbf{{H}}=
\begin{bmatrix}
\mathbf{\widetilde{H}}_1(f_0)&\mathbf{\widetilde{H}}_2(f_0)&\cdots&\mathbf{\widetilde{H}}_{n_x}(f_0)\\
\mathbf{\widetilde{H}}_2(f_0)&\mathbf{\widetilde{H}}_3(f_0)&\cdots&\mathbf{\widetilde{H}}_{n_x+1}(f_0)\\
\vdots&\vdots&\ddots&\vdots\\
\mathbf{\widetilde{H}}_{m_x}(f_0)&\mathbf{\widetilde{H}}_{m_x+1}(f_0)&\cdots&\mathbf{\widetilde{H}}_{m_x}(f_0)
\end{bmatrix}
,
\end{align}
where, 
\begin{align} \label{eq:Hif} 
&\mathbf{\widetilde{H}}_i(f_0)=\\
&\begin{bmatrix}
S(i,1,f_0)&S(i,2,f_0)&\cdots&S(i,n_y,f_0)\\
S(i,2,f_0)&S(i,3,f_0)&\cdots&S(i,n_y+1,f_0)\\
\vdots&\vdots&\ddots&\vdots\\
S(i,m_y,f_0)&S(i,m_y+1,f_0)&\cdots&S(i,N_y,f_0)
\end{bmatrix}
,
\end{align}
$m_x=\text{INT}(N_x/2)+1$ and $n_x=N_x-m_x+1$ are the key parameters to make the block Hankel matrix close to a square matrix. Similarly, $m_y=\text{INT}(N_y/2)+1$ and $n_y=N_y-m_y+1$.  $\text{INT}(\cdot)$ represents the integer part of an argument.

\cite{Zheng2014A} point out that the component with strong energy corresponds to a large singular value, i.e., the strong signals correspond to large singular values. On the contrary, the smaller singular values represent the energy of diffractions and noise. 
It is straightforward to infer that the largest singular values of block Hankel matrix represent the reflections.  If we can obtain the reflection wave $\mathbf{R}$ by extracting the corresponding singular values and vectors, then the diffraction wave $\mathbf{D}$ can be obtained by subtracting $\mathbf{R}$ from the data $\mathbf{S}$,
\begin{equation}\label{eq:dsr}
\mathbf{D}=\mathbf{S}-\mathbf{R}.
\end{equation}

Thus, the problem of diffraction separation is transformed into estimating the reflection, which can be addressed by applying the singular-value decomposition (SVD) algorithm to the block Hankel matrix \cite{Chen2016An}. In particular, truncated singular-value decomposition (TSVD) is usually used to suppress the noise. The TSVD of block Hankel matrix is expressed as
\begin{equation} \label{eq:svd}
\mathbf{\hat{H}}=\mathbf{U}\mathbf{\Sigma}\mathbf{V}^T,
\end{equation}
where $\mathbf{\hat{H}}$ denotes the denoised block Hankel matrix, i.e., the noise can be suppressed by ignoring some smaller singular values \cite{2020Diffraction}.  $\mathbf{\Sigma}=[{\sigma}_1,{\sigma}_2,\cdots,{\sigma}_N]$ is the diagonal singular matrix, $\mathbf{U}=[\mathbf{u}_1,\mathbf{u}_2,\cdots,\mathbf{u}_N]$ and $\mathbf{V}=[\mathbf{v}_1,\mathbf{v}_2,\cdots,\mathbf{v}_N]$ are left and right singular matrices, respectively. They are made up of singular vectors $\mathbf{u}_j$ and $\mathbf{v}_j$. $\sigma_j$ denotes the $j$th singular value, and $N$ denotes the number of singular values.

>>>>

%\documentclass[10pt]{IEEEtran}
%\usepackage{amsmath}
%\usepackage{cite}
%\usepackage{ifthen}
%\usepackage{seg}
%\usepackage{color}
%\usepackage{graphicx}
%\usepackage{subfigure}
%

%\begin{document}

\DeclareRobustCommand{\dlo}[1]{}
\DeclareRobustCommand{\wen}[1]{#1}

\title{3D seismic diffraction separation and imaging using the local rank-reduction method}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\author{Wei Chen, Xingye Liu, Omar M. Saad, Yapo Abol{\'e} Serge Innocent Obou{\'e}, Liuqing Yang, Hang Wang, and Yangkang Chen}
%\thanks{W. Chen is with Cooperative Innovation Center of Unconventional Oil and Gas (Ministry of Education \& Hubei Province), Yangtze University, Wuhan, 430100, China, and Key Laboratory of Exploration Technologies for Oil and Gas Resources of Ministry of Education, Yangtze University, Wuhan, 430100, China}
%\thanks{X. Liu is with Chengdu University of Technology, Dongsanlu, Erxianqiao, Chengdu, 610059, China.}
%\thanks{O.M. Saad is with Seismology Department, National Research Institute of Astronomy and Geophysics (NRIAG), Helwan, 11731, Egypt.}
%\thanks{Y.A.S.I. Oboue and H. Wang are with School of Earth Sciences, Zhejiang University, Hangzhou, 310027, China.}
%\thanks{L. Yang is with State Key Laboratory of Petroleum Resource and Prospecting, China University of Petroleum (Beijing), Beijing, 102200, China.}
%\thanks{Y. Chen is with Bureau of Economic Geology, Jackson School of Geosciences, The University of Texas at Austin, Austin, Texas, 78713-8924, USA.}
%\thanks{The research is supported by the National Natural Science Foundation of China (Grant No. 41804140) and the Open Fund of Cooperative Innovation Center of Unconventional Oil and Gas, Yangtze University (Ministry of Education \& Hubei Province) (Grant No. UOG2021-01).}}

@@

\section{Introduction}
Seismic data contain various useful information that can reflect the geologic features of the subsurface \cite[]{shucai2020,guoyin2018}. In general, reflections are generated by large-scale geologic objects that are continuous and have smooth-reflecting boundaries, and their energy in the seismic records is relatively strong \cite[]{2010Imaging, Peng2018Accurate}. 
Reflections are preserved in conventional seismic exploration workflow, and are the most commonly used signals to detect large-scale geologic structures. 
However, it is difficult to recognize small-scale geologic structures  (faults, small-size scattering points, and reflector edges, etc.) by using reflections. 
These small-scale geologic bodies will scatter the energy that is expressed as diffractions in the seismic records, which are destroyed or removed by the traditional seismic processing techniques.
However, diffractions are helpful to find unconventional reservoirs of oil and gas because they are associated with small-scale discontinuous objects \cite[]{2016Diffraction, 2020Separation}.

Researchers use seismic diffraction to detect subsurface discontinuities, provided that the seismic diffraction is separated from the original data.  
Nevertheless, it is a challenge because the energy of diffractions is weaker compared with the strong energy of reflections, which can even reach several orders of magnitude, i.e., diffractions are usually masked by reflections and noise \cite[]{1994Theory, 2007Post,omar2020geo1}. 
One of the most widely used methods is the path-summation-based method that uses the kinematic differences to separate reflections and diffractions. Based on the coherent summation of diffracted events according to the traveltime, \cite[]{1987A} and \cite[]{1988Imaging} achieve diffraction separation and imaging in common-offset and common-midpoint domains, respectively. 
\cite[]{2004Diffraction} separate and mute the reflection in the pseudodepth domain  by using the difference in properties of waves.
The most known diffraction separation and imaging method is based on a common-reflection-surface method that has been extended by many geophysicists.
 One of the advantages is that this method using several common midpoints to stack based on ray theory \cite[]{2013Recovering}.
\cite[]{2009Diffraction} propose a multi-focusing algorithm to focus diffractions while scattering reflections.
Similarly, \cite[]{2011Common} use the common-reflection-surface approach to achieve diffraction imaging.
\cite[]{2012Separation} achieve diffraction imaging in the migrated dip-angle domain and pointed out that it was more advantageous to separate diffraction from seismic data. \cite[]{2013Diffraction} propose a method that can adaptively choose parameters to fit the diffraction trvaltime and accelerate the computation.
\cite[]{Coimbra2018Enhancement} improve the accuracy of diffraction separation by a finite-offset double-square equation, and extended a fast version without losing accuracy.
\cite[]{2018Fast} develop a robust and fast method to extract the parameters of common-reflection-surface by using the local slope attribute, and it could be used for the diffraction separation. 
\cite[]{2018Common} point out that wavefront attributes of pre-stack seismic data could improve the illumination and are more appropriate for diffraction separation.
\cite[]{20193D} construct a new traveltime difference gather to flatten the diffraction, then muted the upward reflection.
For post-stack seismic data, \cite[]{2019Coherent} uses many wavefront filters to separate reflections and diffractions, and demonstrated its superiority in extracting diffractions masked by strong reflections.

Another effective diffraction separation and imaging method is based on the plane-wave destruction (PWD) proposed by \cite[]{Sergey2002Applications} and  \cite[]{2007Post}. Traveltimes of diffractions are quasi-hyperbolic-shaped while those of reflections are laterally continuous in the constant plane-wave slope sections. Thus, reflections can be predicted from neighbor traces by using a prediction-error filter. Then, the predicted reflection is suppressed and the diffraction is extracted.  After that, diffractions are migrated by selecting an optimal velocity. That is to say, the PWD-based method can simultaneously achieve velocity estimation and diffraction imaging \cite[]{2011Azimuthally}. 
\cite[]{2013Coal} utilize the PWD-based diffraction imaging method to detect the fractures in coals.
\cite[]{Binzhong2018seismic} propose a local moving-average-error-filter to achieve diffraction separation and applied it to coal structures detection. 
\cite[]{2018l1} developed an improved PWD-based method via wavelet transform and L1-norm, and achieved accurate diffraction separation.
The PWD-based method depends on the local slopes, thus a wide range of approaches are developed to accurately estimate the local slopes, such as \cite[]{2008Separation, 2010Nonlinear, 2018Dip}.
\cite[]{2020Least} achieve the diffraction separation by establishing an inversion objection and combining the PWD algorithm with other advanced techniques, such as Kirchhoff modeling, and path-summation-based methods.

Most diffraction separation and imaging methods only take kinematic features of diffractions and reflections into consideration, and the dynamic differences between diffractions and reflections are ignored. However, the dynamic discrepancy is significant to diffraction imaging \cite[]{2015Diffraction, 20193D, 2020Diffraction}.
\cite[]{2014Diffraction} update migration velocities according to shapes of reflections and diffractions in the migrated gathers by picking the angles related to reflections in the shot and opening-angle gathers, improving the accuracy.
\cite[]{2015Diffraction} develop a least-squares fitting method by using double exponential functions to analyze the amplitude function of diffracted waves and improve the performance of diffraction separation.
\cite[]{2016Imaging} develop a three-step diffraction separation and imaging approach based on wave-equation. It is computationally fast and resistant to random noise.
Based on the fact that the dip direction of reflectors could not be observed in both left- and right-downgoing propagating waves, \cite[]{2019Efficient} develop an efficient wave-equation-based approach for diffraction separation.
In frequency-wavenumber domain, \cite[]{2020Inspiration} use a double square root equation to simulate diffractions, and separated diffractions by using dip frequency filters.
To improve the image quality of diffractions, \cite[]{2020Periodic} propose a periodic plane-wave least-squares reverse time migration method.
\cite[]{2021Diffraction} applies a diffraction-angle filtering method to the virtual source in the acoustic full-waveform inversion to separate energy of reflections and transmissions. Diffractions can be separated by designing many types of diffraction-angle filters.

Consequently, both kinematic and dynamic differences between reflections and diffractions are important for seismic diffraction separation and imaging. Jointly using kinematic and dynamic properties of wave fields is helpful for diffraction separation and imaging \cite[]{2010Separation}. 
\cite[]{2020Separation} propose a 2D diffraction separation method based on a local rank-reduction algorithm, in which the rank is optimized by an adaptive selection approach.  \cite[]{2020Diffraction} utilize the multichannel singular-spectrum  analysis (MSSA) for diffraction separation and imaging. Their implementation of the MSSA method does not consider the strong non-stationarity of seismic data and thus does not use local windows for the diffraction separation, which is referred to as a global rank-reduction (GRR) method. In the post-stack domain, these two methods make full use of kinematic and dynamic differences between reflections and diffractions. Besides, these methods were tested on 2D seismic data. For 3D data, the complexity would increase and a unique rank value is unsatisfactory for accurate diffraction separation and imaging. 
Therefore, we present a novel method based on the local rank-reduction (LRR) method for 3D seismic diffraction separation. Both kinematic and dynamic properties are considered. The rank-reduction filter is carried out in each local window to enhance the performance \cite[]{shaohuan2017gji,yangkang2017ieee,yangkang2019nc}. In this case, specifying an appropriate rank for different windows is necessary to make the result more promising. Due to the rank inconsistency,  we propose an automatic selection approach to locally optimize the rank at different locations, which is not essential in the global strategy.

@@

%\begin{figure}[htb!]
%\centering
%\includegraphics[width=\columnwidth]{Fig/demosigma}
%\caption{Singular value distribution. (a) 3D reflection data containing three plane waves. (b) 3D seismic data containing reflections, diffractions and random noise. (c) 2D section extracted from (a). (d) 2D section extracted from (b). (e) Singular values distribution of the Hankel matrix in the 20 Hz slice of the data shown in (a).  (f) Singular values distribution of the data shown in (b). }
%\label{fig:demosigma}
%\end{figure}
\inputdir{./}
\plot{demosigma}{width=\columnwidth}{Singular value distribution. (a) 3D reflection data containing three plane waves. (b) 3D seismic data containing reflections, diffractions and random noise. (c) 2D section extracted from (a). (d) 2D section extracted from (b). (e) Singular values distribution of the Hankel matrix in the 20 Hz slice of the data shown in (a).  (f) Singular values distribution of the data shown in (b). }
%\begin{figure}[htb!]
% \centering
% \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/refl}
%   \label{fig:refl}}
% \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/diffr}
%   \label{fig:diffr}}   
% \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/syn}
%   \label{fig:syn}}    
%\caption{Benchmark dataset with groundtruth reflections and diffractions. (a) Reflection data. (b) Diffraction data. (c) Synthetic data. }
%\label{fig:refl,diffr,syn}
%\end{figure}
%
%\begin{figure}[htb!]
% \centering
% \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-pwd-0}
%   \label{fig:s-pwd-0}}
% \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-grr-0}
%   \label{fig:s-grr-0}}   
% \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-lrra-0}
%   \label{fig:s-lrra-0}}    
%\caption{Separated reflections using (a) PWD method, (b) global rank-reduction method, and (c) proposed method.}
%\label{fig:s-pwd-0,s-grr-0,s-lrra-0}
%\end{figure}
%
%\begin{figure}[htb!]
% \centering
% \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-pwd-n-0}
%   \label{fig:s-pwd-n-0}}
% \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-grr-n-0}
%   \label{fig:s-grr-n-0}}   
% \subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-lrra-n-0}
%   \label{fig:s-lrra-n-0}}    
%\caption{Separated diffractions using (a) PWD method (SNR=6.52 dB), (b) global rank-reduction method (SNR=1.42 dB), and (c) proposed method (SNR=9.14 dB).}
%\label{fig:s-pwd-n-0,s-grr-n-0,s-lrra-n-0}
%\end{figure}
%
%\begin{figure}[htb!]
%\centering
%\subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-pwd-e-0}
%  \label{fig:s-pwd-e-0}}
%\subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-grr-e-0}
%  \label{fig:s-grr-e-0}}   
%\subfigure[]{\includegraphics[width=0.31\columnwidth]{plot/Fig/s-lrra-e-0}
%  \label{fig:s-lrra-e-0}}    
%\caption{\new{Error of separated diffractions using (a) PWD method, (b) global rank-reduction method, and (c) proposed method.} \color{black}{}}
%\label{fig:s-pwd-e-0,s-grr-e-0,s-lrra-e-0}
%\end{figure}

\inputdir{syn3d}
\multiplot{3}{refll,diffrr,syn3d2}{width=0.45\columnwidth}{Benchmark dataset with groundtruth reflections and diffractions. (a) Reflection data. (b) Diffraction data. (c) Synthetic data. }
\multiplot{3}{s-pwd-0,s-grr-0,s-lrra-0}{width=0.45\columnwidth}{Separated reflections using (a) PWD method, (b) global rank-reduction method, and (c) proposed method.}
\multiplot{3}{s-pwd-n-0,s-grr-n-0,s-lrra-n-0}{width=0.45\columnwidth}{Separated reflections using (a) PWD method (SNR=6.52 dB), (b) global rank-reduction method (SNR=1.42 dB), and (c) proposed method (SNR=9.14 dB).}
\multiplot{3}{s-pwd-e-0,s-grr-e-0,s-lrra-e-0}{width=0.45\columnwidth}{Error of separated diffractions using (a) PWD method, (b) global rank-reduction method, and (c) proposed method.}

\section*{THEORY AND Methods}
Diffraction separation can be regarded as a problem of seismic data reconstruction, in which diffractions are treated as effective signals while reflections are treated as noise.  In the post-stack seismic data, reflections appear as locally linear events, while diffractions behave as hyperbolic events.  The rank-reduction method can utilize the kinematic differences between diffractions and reflections to separate different components. Besides, the energy of diffractions is much weaker than that of reflections, thereby corresponding to different singular values and singular vectors. The distributions of singular values for diffractions and reflections are assumed to be disparate in the singular-value spectrum. Therefore, the contribution of the last few singular values is the smallest, while the first few singular values contribute more to reconstructing the data. The rank-reduction method also considers the energy discrepancy, i.e., the dynamic differences, to separate weak diffractions and strong reflections \cite[]{2010Separation, 2016Diffraction}.

\subsection{Local rank-reduction method}
Let a 3D seismic data be denoted by $S(x, y, t)$, $x=1,\cdots, N_x$, $y=1,\cdots, N_y$, and $t=1,\cdots, N_t$, where $N_x$, $N_y$ and $N_t$ represent the numbers of samples in spatial and time axes, respectively. The rank-reduction method \cite[]{2011Simultaneous,weilin2016dmssa} is implemented in the frequency domain, thus the first step is to obtain the corresponding frequency-domain data $S(x, y,f )$  by forward Fourier transform $\mathcal{F}$. For a given frequency $f_0$, we construct the block Hankel matrix $\mathbf{{H}}$ by mapping the frequency-domain data into several Hankel matrices $\mathbf{\widetilde{H}}_i(f_0)$ through the Hankelization operator $\mathcal{H}$. Then, we can obtain,
\begin{align} \label{eq:B} 
\mathbf{{H}}=
\begin{bmatrix}
\mathbf{\widetilde{H}}_1(f_0)&\mathbf{\widetilde{H}}_2(f_0)&\cdots&\mathbf{\widetilde{H}}_{n_x}(f_0)\\
\mathbf{\widetilde{H}}_2(f_0)&\mathbf{\widetilde{H}}_3(f_0)&\cdots&\mathbf{\widetilde{H}}_{n_x+1}(f_0)\\
\vdots&\vdots&\ddots&\vdots\\
\mathbf{\widetilde{H}}_{m_x}(f_0)&\mathbf{\widetilde{H}}_{m_x+1}(f_0)&\cdots&\mathbf{\widetilde{H}}_{m_x}(f_0)
\end{bmatrix}
,
\end{align}
where, 
\begin{align} \label{eq:Hif} 
&\mathbf{\widetilde{H}}_i(f_0)=\\
&\begin{bmatrix}
S(i,1,f_0)&S(i,2,f_0)&\cdots&S(i,n_y,f_0)\\
S(i,2,f_0)&S(i,3,f_0)&\cdots&S(i,n_y+1,f_0)\\
\vdots&\vdots&\ddots&\vdots\\
S(i,m_y,f_0)&S(i,m_y+1,f_0)&\cdots&S(i,N_y,f_0)
\end{bmatrix}
,
\end{align}
$m_x=\text{INT}(N_x/2)+1$ and $n_x=N_x-m_x+1$ are the key parameters to make the block Hankel matrix close to a square matrix. Similarly, $m_y=\text{INT}(N_y/2)+1$ and $n_y=N_y-m_y+1$.  $\text{INT}(\cdot)$ represents the integer part of an argument.

\cite[]{Zheng2014A} point out that the component with strong energy corresponds to a large singular value, i.e., the strong signals correspond to large singular values. On the contrary, the smaller singular values represent the energy of diffractions and noise. 
It is straightforward to infer that the largest singular values of block Hankel matrix represent the reflections.  If we can obtain the reflection wave $\mathbf{R}$ by extracting the corresponding singular values and vectors, then the diffraction wave $\mathbf{D}$ can be obtained by subtracting $\mathbf{R}$ from the data $\mathbf{S}$,
\begin{equation}\label{eq:dsr}
\mathbf{D}=\mathbf{S}-\mathbf{R}.
\end{equation}
